Draft Manuscript

### 2022-09-23 Prereg_and_manuscript/manuscript.Rmd
Added resutls and discussion to manuscript

csl: https://www.zotero.org/styles/ecology-letters
bibliography: references.bib
---

@@

@@

```{r include=TRUE}
# creating directory to store data
dir.create("data")

@@

```{r}
# Loading data tables ----
X_mat <- read.csv(here::here("data","Matrice_X.csv"),
                  sep = ";") # explanatory variables - see metadata
Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                  sep = ";") # response variables (mean ericaceous cover, in %)
clim <- read.csv(here::here("data","Climate_value.csv"),
                 sep = ";") # numerical codes used for clim vars in X_mat

@@

for (i in 1:dim(X_mat)[1]){
  most_abun <-  as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), k=1, descending = T,index.return = T)) # returning the index of the column with the most abundant tree species for each row
  sec_abun <- as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), k=2, descending = T,index.return = T)) # returning index of second most abundant species
  dominant_sp <- c(most_abun, sec_abun)+1 # adding 1 to account for the first column which was left out of vector for previous function
  dominant_sp <- colnames(X_mat)[dominant_sp] # creating character vector containing column names of most dominant species

@@

```{r}
# Classifying as high or low precipitation ----
# PRECU is the annual rainfall variable

@@

```{r}
# Combining and classifying as forest type - rainfall index
X_mat <- 
  X_mat %>% 
  unite(., rf_index, c("RainFall","Forest_Type"), sep = "_", remove =F, na.rm = FALSE)

@@

```{r}
# loading in prepared data ----
  X_mat <- read.csv(here::here("data","X_mat_processed.csv"))
  X_mat <- X_mat %>% 
    select(CLE, Forest_Type, PRECU, RainFall,rf_index)

@@

  # For Chamaedaphne calyculata
  CAL <- data.frame(Y_mat$V_CAL,X_mat$PRECU, X_mat$rf_index) # creating dataframe with only this species
  colnames(CAL) <- c("V_CAL","PRECU","rf_index") # renaming columns 
  ggplot(data=CAL, aes(x=rf_index, y=V_CAL)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=CAL, aes(x=rf_index, y=log(V_CAL))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

  # For Rhododendron groenlandicum
  RHG <- data.frame(Y_mat$V_RHG,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
  colnames(RHG) <- c("V_RHG","PRECU","rf_index") # renaming columns 
  ggplot(data=RHG, aes(x=rf_index, y=V_RHG)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=RHG, aes(x=rf_index, y=log(V_RHG))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

## For Kalmia angustifolia
  # Plotting: 
  KAA <- data.frame(Y_mat$V_KAA,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
  colnames(KAA) <- c("V_KAA","PRECU","rf_index") # renaming columns

  ggplot(data=KAA, aes(x=rf_index, y=V_KAA)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=KAA, aes(x=rf_index, y=log(V_KAA))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

## For Vaccinium sp. 
    VAAM <- data.frame(Y_mat$V_VAAM,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
    colnames(VAAM) <- c("V_VAAM","PRECU","rf_index") # renaming columns

    ggplot(data=VAAM, aes(x=rf_index, y=V_VAAM)) + # plotting only this species with the rainfall-forest type index in box plots
      geom_boxplot()

    ggplot(data=VAAM, aes(x=rf_index, y=log(V_VAAM))) + # log transforming the y axis for easier pattern detection
      geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

@@

```{r}
## Plotting altogether in facets 
    # creating new variable to denote species to make table with all species
    CAL <- CAL %>% 
      add_column("species" = "CAL") %>%  # adding column with species names
      rename(cover ="V_CAL") # renaming percent cover column to represent all species

@@

    plot_cover <- rbind(CAL,RHG, KAA, VAAM) # combining all species by rows into one table

    ## creating ggplot object
    p <- ggplot(data= plot_cover, aes(x=rf_index, y=log(cover))) + # log transforming the y axis for easier pattern detection
      geom_boxplot()+
      xlab("Rainfall-Forest Type Index (RF)") + # changing x axis labe
      ylab("Log (% cover)")+ # changing y axis label
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5), # changing plot themes for angled text and no gridlines
            panel.grid = element_blank())

@@

```{r}
 # We want to know is there any significant difference between percent over of plants in different rainfall and forest type conditions

@@

# Results

\# CAL significant differences (p\< 0.05): H_d vs H_c ; H_m vs H_d ; L_c vs H_d ; L_m vs H_c ; L_m vs L_c

@@

\# RHGsignificant differences (P\<0.05): all types with H_c; H_d vs H_m ; H_d vs H_n ; H_d vs L_c ; L_c vs L_d ; L_c vs L_m ; L_c vs L_n

@@

# Discussion

@@

@@

# Tables

**Table 1.** Mean body mass of penguins on different islands over time.

Note: for the word version the `kable` does not output correctly. You could try using `flextable` package instead: <https://taehoonh.me/content/post/alternative-to-kable-function-when-knitting-to-ms-word.html>

\`\`\`{#{r table_01, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}

# without the "hold_position" the table ends up at the top of the page

    # Figure Captions

    **Figure 1**. Pretty coloured dots about penguins

    **Figure 2**. Wow, even prettier plot about penguins that shows stuff

    # Figures

    ![]()

    Figure 1.

    ```{#{r figure_02, echo=FALSE, message=FALSE, warning=FALSE}

Figure 2.

Figure 3.

`{#{r file="../scripts/figure_03.R"} # this code chunk calls an external script to generate the plot. If you # want to load or run code not for a plot you could also use the following. # Note that it does not seem to work for plots. # source("../scripts/figure_03.R", local = knitr::knit_global())`

# Appendices

```{=html}
<!-- You might want to look into Bookdown if you want to do anything fancier
with your manuscript. Bookdown allows for automatic Table and Figure numbering
as well as cross referencing. That is you can create links between 'Fig 1' in
your manuscript that links to the actual figure and automatically the
the figure number as you add or remove them from the document.
I add this here because Appendices can require special handling by some 
journals and Bookdown might allow you to do that automagically -->
```

>>>>

csl: https://www.zotero.org/styles/ecology-letters
bibliography: [references.bib,grateful-refs.bib]
---

@@

# remotes::install_github("Pakillo/grateful")
# library(grateful)
# get_pkgs_info()
```

@@

```{r eval=FALSE, warning=FALSE, include=FALSE}
# creating directory to store data
dir.create("data")

@@

```{r, include=TRUE,echo=TRUE}
# Loading data tables ----
X_mat <- read.csv(here::here("data","Matrice_X.csv"),
                  sep = ";") # explanatory variables - see metadata
Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                  sep = ";") # response variables (mean ericaceous cover, in %)
clim <- read.csv(here::here("data","Climate_value.csv"),
                 sep = ";") # numerical codes used for clim vars in X_mat

@@

for (i in 1:dim(X_mat)[1]){
  most_abun <-  as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), 
                                      k=1, descending = T,index.return = T)) 
  # returning the index of the column with the most abundant tree species for each row
  sec_abun <- as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), 
                                    k=2, descending = T,index.return = T)) 
  # returning index of second most abundant species
  dominant_sp <- c(most_abun, sec_abun)+1 # adding 1 to account for the 
  # first column which was left out of vector for previous function
  dominant_sp <- colnames(X_mat)[dominant_sp] # creating character vector containing 
  # column names of most dominant species

@@

```{r,include=TRUE,echo=TRUE}
# Classifying as high or low precipitation ----
# PRECU is the annual rainfall variable

@@

```{r,include=TRUE,echo=TRUE}
# Combining and classifying as forest type - rainfall index
X_mat <- 
  X_mat %>% 
  unite(., rf_index, c("RainFall","Forest_Type"), sep = "_", remove =F, na.rm = FALSE)

@@

```{r, include=TRUE,echo=TRUE}
# loading in prepared data ----
  X_mat <- read.csv(here::here("data","X_mat_processed.csv"))
  X_mat <- X_mat %>% 
    select(CLE, Forest_Type, PRECU, RainFall,rf_index)

@@

  # For Chamaedaphne calyculata
  # creating dataframe with only this species
  CAL <- data.frame(Y_mat$V_CAL,X_mat$PRECU, X_mat$rf_index) 
  colnames(CAL) <- c("V_CAL","PRECU","rf_index") # renaming columns 
  ggplot(data=CAL, aes(x=rf_index, y=V_CAL)) + 
    geom_boxplot() # plotting only this species with the rainfall-forest
    # type index in box plots

  CAL_plot <- ggplot(data=CAL, aes(x=rf_index, y=log(V_CAL))) + # log transforming the y axis 
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

  # Repeating for Rhododendron groenlandicum
  RHG <- data.frame(Y_mat$V_RHG,X_mat$PRECU,X_mat$rf_index) 
  colnames(RHG) <- c("V_RHG","PRECU","rf_index") 
  RHG_plot <- ggplot(data=RHG, aes(x=rf_index, y=V_RHG)) + 
    geom_boxplot()

  RHG_plot <- ggplot(data=RHG, aes(x=rf_index, y=log(V_RHG))) + # log transforming the y axis 
    geom_boxplot()

## Repeating for Kalmia angustifolia
  # Plotting: 
  KAA <- data.frame(Y_mat$V_KAA,X_mat$PRECU,X_mat$rf_index) 
  colnames(KAA) <- c("V_KAA","PRECU","rf_index") 
  KAA_plot <- ggplot(data=KAA, aes(x=rf_index, y=V_KAA)) + 
    geom_boxplot()

  KAA_plot <- ggplot(data=KAA, aes(x=rf_index, y=log(V_KAA))) + # log transforming the y axis 
    geom_boxplot()

## Repeating for Vaccinium sp. 
    VAAM <- data.frame(Y_mat$V_VAAM,X_mat$PRECU,X_mat$rf_index) 
    colnames(VAAM) <- c("V_VAAM","PRECU","rf_index")

   VAAM_plot <-ggplot(data=VAAM, aes(x=rf_index, y=V_VAAM)) +
      geom_boxplot()

    VAAM_plot <- ggplot(data=VAAM, aes(x=rf_index, y=log(V_VAAM))) + # log transforming the y axis 
      geom_boxplot()

@@

```{r,echo=FALSE, include=FALSE}
## Plotting altogether in facets 
    # creating new variable to denote species to make table with all species
    CAL <- CAL %>% 
      add_column("species" = "CAL") %>%  # adding column with species names
      rename(cover ="V_CAL") # renaming percent cover column to represent all species

@@

    plot_cover <- rbind(CAL,RHG, KAA, VAAM) # combining all species by rows into table

    ## creating ggplot object
    p <- ggplot(data= plot_cover, aes(x=rf_index, y=log(cover))) + # log transforming 
      geom_boxplot()+
      xlab("Rainfall-Forest Type Index (RF)") + # changing x axis labe
      ylab("Log (% cover)")+ # changing y axis label
      # changing plot themes for angled text and no gridlines
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5), 
            panel.grid = element_blank())

@@

```{r,echo=FALSE, include=FALSE}
 # We want to know is there any significant difference between percent over of plants in different rainfall and forest type conditions

@@

# Results & Discussion

```{r}
p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)
```

**Figure 1.** Box plots of *C. calyculata*, *K. angustifolia*, *R. groenlandicum*, and *Vaccinium sp*. percent cover in plots that fall in each of the rainfall-forest type categories. The thick horizontal lines represent the median for each category. The y axis was log-transformed for display purposes.

**Table 1.** Mean % cover for *C. calyculata* in different rainfall forest type categories.

```{r}
CAL_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *C. calyculata,* the only rainfall forest type category that had median percent cover above 0 was high rainfall coniferous forests. This supports the prediction that this species would have a greater preference for this category possibly due to the higher acidity and moisture provided.

**Table 2.** Mean % cover for *R. groenlandicum* in different rainfall forest type categories.

```{r}
RHG_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *R. groenlandicum*, the order of my predictions also matched those observed. H_c had significantly higher percent cover than any other rf type (p\<0.05), followed by L-c, H_m, L_m H_d and L_d. This may indicate that *R. groenlandicum* indeed prefers more acidic soils created by higher rainfall, but mroe importantly coniferous forests. However, this is confounded with other factors that may drive its preference for coniferous habitat such as other soil properties associated with conifer needles and light availability.

**Table 3.** Mean % cover for *K. angustifolium* in different rainfall forest type categories.

```{r}
KAA_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *K. angustifolium* the patterns in abundance that I predicted (L-c, L-m, L-d, H-c, H-m, H-d) were not supported by those observed. The category with the highest median percent cover was non-forested area however this was not significantly different from abundance in high rainfall coniferous forests (p\>0.05). This may indicate that light-availability is a more important determinant of its distribution. Additionally, non-forested plots may often be boggy areas with acidic soils from sphagnum growth, so perhaps this species' distribution is highly driven by soil acidity, so much so that the acidity produced by conifer needles and high rainfall is not sufficient.

```{r}
VAAM_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

Finally, for *Vaccinium sp.*

@@

@@

Analysis to test if these species distributions are driven partly by soil acidity could ideally be performed using direct measures of soil acidity at the plots rather than the proxy variables used here. Upon future visits to the sites, these measurements could be taken to construct models and compare preferences between species to better predict what drives their distribution patterns and regenerative ability in logging and fire disturbed habitats.

@@

```{r results = 'asis'}
grateful::cite_packages(output = "paragraph")
```

@@

# Appendix

**Table A1.**
